---
title: "hoeffdingtest"
output: html_document
date: "2024-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(Hmisc)
library(tidyverse)
library(R.matlab)
```

## Import data into R and perform analysis

Now, we'll import the data into R, reshape it back to 3D, calculate the means, and perform the Hoeffding independence test using the Hmisc library.

```{r}
# Import data from .mat file
mat_data <- readMat('runtime_data.mat')
runtime_array <- mat_data$runtime.list

# Calculate means
means <- apply(runtime_array, c(1, 2), mean)
stds <- apply(runtime_array, c(1, 2), sd)

# Recreate the combinations_indices
combinations_indices <- lapply(0:3, function(i) c(i, 0, i, 1, i, 49, i, 99))
combinations_indices <- matrix(unlist(combinations_indices), ncol=2, byrow=TRUE)

# Function to perform Hoeffding test
perform_hoeffding_test <- function(c_1, c_2) {
  r_1 <- (runtime_array[, c_1[1] + 1, c_1[2] + 1] - means[, c_1[1] + 1])/(stds[, c_1[1] +1])
  r_2 <-  (runtime_array[, c_2[1] + 1, c_2[2] + 1] - means[, c_2[1] + 1])/stds[, c_2[1] + 1]
  result <- hoeffd(r_1, r_2)
  return(list(D = result$D, p_value = result$P))
}

# Perform pairwise comparisons
n_combinations <- nrow(combinations_indices)
results_matrix <- matrix(NA, nrow = n_combinations, ncol = n_combinations)

for (i in 1:(n_combinations - 1)) {
  for (j in (i + 1):n_combinations) {
    c_1 <- combinations_indices[i, ]
    c_2 <- combinations_indices[j, ]
    
    test_result <- perform_hoeffding_test(c_1, c_2)
    results_matrix[i, j] <- test_result$p_value[1, 2]
    results_matrix[j, i] <- test_result$p_value[1, 2]
    
    cat("Comparison:", paste(c_1, collapse=","), paste(c_2, collapse=","), 
        "D:", test_result$D, "p-value:", test_result$p_value[1,2], "\n")
  }
}

# Set diagonal to 1 (self-comparison)
diag(results_matrix) <- 1

# Export P-value matrix as .mat file
writeMat("p_value_matrix.mat", p_value_matrix = results_matrix)

cat("\nP-value matrix exported to 'p_value_matrix.mat'\n")
```

## Conclusion

This analysis performs the Hoeffding independence test using the `hoeffd` function from the Hmisc library in R, maintaining the 3D structure of the input data. It calculates means directly in R and provides p-values for each pairwise comparison. The P-value matrix is exported to a CSV file for further analysis in Python.

The results are summarized and visualized to give an overview of the independence relationships in the data. The P-value matrix can be found in the 'p_value_matrix.csv' file for use in your Python analysis code.
