---
title: "hoeffdingtest"
output: html_document
date: "2024-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
library(Hmisc)
library(tidyverse)
library(R.matlab)
```

## Import data into R and perform analysis

Now, we'll import the data into R, reshape it back to 3D, calculate the means, and perform the Hoeffding independence test using the Hmisc library.

```{r}
# Import data from .mat file
mat_data <- readMat('runtime_data.mat')
runtime_array <- mat_data$runtime.list

# Calculate means
means <- apply(runtime_array, c(1, 2), mean)
stds <- apply(runtime_array, c(1, 2), sd)

# Recreate the combinations_indices
combinations_indices <- lapply(0:3, function(i) c(i, 0, i, 1, i, 49, i, 99))
combinations_indices <- matrix(unlist(combinations_indices), ncol=2, byrow=TRUE)

# Function to perform Hoeffding test
perform_hoeffding_test <- function(c_1, c_2) {
  #r_1 <- log(runtime_array[, c_1[1] + 1, c_1[2] + 1]) - log(means[, c_1[1] + 1])
  #r_2 <-  log(runtime_array[, c_2[1] + 1, c_2[2] + 1]) - log(means[, c_2[1] + 1])
  r_1 <- (runtime_array[, c_1[1] +1, c_1[2] + 1] - means[, c_1[1]+1])/stds[, c_1[1] +1]
  r_2 <- (runtime_array[, c_2[1] +1, c_2[2] +1] - means[, c_2[1] + 1])/stds[, c_2[1] + 1]
  result <- hoeffd(r_1, r_2)
  return(list(D = result$D, p_value = result$P))
}

# Perform pairwise comparisons
n_combinations <- nrow(combinations_indices)
results_matrix <- matrix(NA, nrow = n_combinations, ncol = n_combinations)
smaller_i <- 0
for (i in 1:(n_combinations - 1)) {
  for (j in (i + 1):n_combinations) {
    c_1 <- combinations_indices[i, ]
    c_2 <- combinations_indices[j, ]
    
    test_result <- perform_hoeffding_test(c_1, c_2)
    results_matrix[i, j] <- test_result$p_value[1, 2]
    results_matrix[j, i] <- test_result$p_value[1, 2]
    if(test_result$p_value[1, 2] < 0.05){
      smaller_i <- smaller_i + 1
    }
    
    cat("Comparison:", paste(c_1, collapse=","), paste(c_2, collapse=","), 
        "D:", test_result$D, "p-value:", test_result$p_value[1,2], "\n")
  }
}

# Set diagonal to 1 (self-comparison)
diag(results_matrix) <- 1

# Export P-value matrix as .mat file
writeMat("p_value_matrix.mat", p_value_matrix = results_matrix)

cat("\nP-value matrix exported to 'p_value_matrix.mat'\n")
```

## Conclusion

This analysis performs the Hoeffding independence test using the `hoeffd` function from the Hmisc library in R, maintaining the 3D structure of the input data. It calculates means directly in R and provides p-values for each pairwise comparison. The P-value matrix is exported to a CSV file for further analysis in Python.

The results are summarized and visualized to give an overview of the independence relationships in the data. The P-value matrix can be found in the 'p_value_matrix.csv' file for use in your Python analysis code.

```{r}

# Generate all combinations of the first index (0 to 3) and second index (0 to 99)
combinations_indices <- expand.grid(0:3, 0:99)

# Convert the result into a matrix, where each row is a combination of (i, j)
combinations_indices <- as.matrix(combinations_indices)

n_combinations <- nrow(combinations_indices)
tests <- 0
smaller_i <- 0
p_values <- numeric()
for (i in 1:(n_combinations - 1)) {
  for (j in (i + 1):n_combinations) {
    c_1 <- combinations_indices[i, ]
    c_2 <- combinations_indices[j, ]
    
    test_result <- perform_hoeffding_test(c_1, c_2)
    tests <- tests + 1
    p_value <- test_result$p_value[1, 2]
    p_values <- c(p_values, p_value)
    if(p_value < 0.05){
      smaller_i <- smaller_i + 1
    }
  }
}
# At the end, print out the quantiles of the p-value distribution
quantiles <- quantile(p_values, probs = c(0.25, 0.5, 0.75, 0.95, 0.99))
print(quantiles)
print(head(p_values))
```
